2-factor-Hull-White\_model\_calibration

### 2020-04-09 Submission1.tex
Introduction finished

    \usepackage{graphicx} %package to manage images
    \graphicspath{ {./images/} }
    \usepackage{wrapfig}

@@

\section{Introduction
}
The pricing calculations in this paper concern a European up-and-out call option. The up-and-out option is a type barrier option that is worthless if the barrier price is reached or exceeded, but otherwise behaves exactly as a standard European call option. There are four main variants of barrier options, two of which are cancelled by reaching the barrier price (up-and-out and down-and-out) and two of which are only valid if the barrier is reached (up-and-in and down-and-in).

The standard European call option payoff formula can be applied with the addition of a check to determine if the barrier price was reached or exceeded, in which case the payoff is zero for the up-and-out option. The same approach could be applied with minor adjustments to price down-and-out, down-and-in or up-and-in barrier options.

Possible paths for the up-and-out barrier option payoff would look similar to the chart below, where two of the paths have crossed the barrier and resulted in zero payoff, but the other three have not crossed the barrier and behave like a vanilla option.

\begin{figure}
    \centering
    \caption{Simulated Payoffs for an Up-and-out Barrier Option}
    \includegraphics{images/barrier-example.png}
\end{figure}

\section{Simulated Paths for the Underlying Share and Counterparty Firm
}

 In order to simulate paths for the underlying share and the firm, a matrix of correlated prices is generated based on the requirement for a correlation of 0.2.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{4}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{k}{def} \PY{n+nf}{share\PYZus{}path}\PY{p}{(}\PY{n}{S\PYZus{}0}\PY{p}{,} \PY{n}{risk\PYZus{}free\PYZus{}rate}\PY{p}{,} \PY{n}{sigma}\PY{p}{,} \PY{n}{Z}\PY{p}{,} \PY{n}{dT}\PY{p}{)}\PY{p}{:}
    \PY{k}{return} \PY{n}{S\PYZus{}0}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{exp}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{cumsum}\PY{p}{(}\PY{p}{(}\PY{n}{risk\PYZus{}free\PYZus{}rate}\PY{o}{\PYZhy{}}\PY{n}{sigma}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{o}{/}\PY{l+m+mi}{2}\PY{p}{)}\PY{o}{*}\PY{n}{dT} \PY{o}{+} \PY{n}{sigma}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{sqrt}\PY{p}{(}\PY{n}{dT}\PY{p}{)}\PY{o}{*}\PY{n}{Z}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{)}\PY{p}{)}

\PY{k}{def} \PY{n+nf}{generate\PYZus{}share\PYZus{}and\PYZus{}firm\PYZus{}price}\PY{p}{(}\PY{n}{S0}\PY{p}{,} \PY{n}{v\PYZus{}0}\PY{p}{,} \PY{n}{risk\PYZus{}free}\PY{p}{,} \PY{n}{sigma}\PY{p}{,} \PY{n}{sigma\PYZus{}firm}\PY{p}{,} \PY{n}{corr}\PY{p}{,} \PY{n}{T}\PY{p}{,} \PY{n}{sample\PYZus{}size} \PY{o}{=} \PY{l+m+mi}{10}\PY{p}{,} \PY{n}{timesteps} \PY{o}{=} \PY{l+m+mi}{12}\PY{p}{)}\PY{p}{:}
    \PY{n}{corr\PYZus{}matrix} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{p}{[}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,} \PY{n}{corr}\PY{p}{]}\PY{p}{,} \PY{p}{[}\PY{n}{corr}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{]}\PY{p}{]}\PY{p}{)}
    \PY{n}{norm\PYZus{}matrix} \PY{o}{=} \PY{n}{stats}\PY{o}{.}\PY{n}{norm}\PY{o}{.}\PY{n}{rvs}\PY{p}{(}\PY{n}{size} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{p}{[}\PY{n}{sample\PYZus{}size}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{,} \PY{n}{timesteps}\PY{p}{]}\PY{p}{)}\PY{p}{)}
    \PY{n}{corr\PYZus{}norm\PYZus{}matrix} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{matmul}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{linalg}\PY{o}{.}\PY{n}{cholesky}\PY{p}{(}\PY{n}{corr\PYZus{}matrix}\PY{p}{)}\PY{p}{,} \PY{n}{norm\PYZus{}matrix}\PY{p}{)}

    \PY{n}{share\PYZus{}price\PYZus{}path} \PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{share\PYZus{}path}\PY{p}{(}\PY{n}{S0}\PY{p}{,} \PY{n}{risk\PYZus{}free}\PY{p}{,} \PY{n}{sigma}\PY{p}{,} \PY{n}{Z}\PY{o}{=}\PY{n}{corr\PYZus{}norm\PYZus{}matrix}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{l+m+mi}{0}\PY{p}{,}\PY{p}{]}\PY{p}{,} \PY{n}{dT}\PY{o}{=}\PY{l+m+mi}{1}\PY{o}{/}\PY{n}{timesteps}\PY{p}{)}\PY{p}{)}
    \PY{n}{share\PYZus{}price\PYZus{}path} \PY{o}{=} \PY{n}{share\PYZus{}price\PYZus{}path}\PY{o}{.}\PY{n}{transpose}\PY{p}{(}\PY{p}{)}

    \PY{n}{firm\PYZus{}price\PYZus{}path} \PY{o}{=} \PY{n}{pd}\PY{o}{.}\PY{n}{DataFrame}\PY{p}{(}\PY{n}{share\PYZus{}path}\PY{p}{(}\PY{n}{v\PYZus{}0}\PY{p}{,} \PY{n}{risk\PYZus{}free}\PY{p}{,} \PY{n}{sigma\PYZus{}firm}\PY{p}{,} \PY{n}{Z}\PY{o}{=}\PY{n}{corr\PYZus{}norm\PYZus{}matrix}\PY{p}{[}\PY{p}{:}\PY{p}{,}\PY{l+m+mi}{1}\PY{p}{,}\PY{p}{]}\PY{p}{,} \PY{n}{dT}\PY{o}{=}\PY{l+m+mi}{1}\PY{o}{/}\PY{n}{timesteps}\PY{p}{)}\PY{p}{)}
    \PY{n}{firm\PYZus{}price\PYZus{}path} \PY{o}{=} \PY{n}{firm\PYZus{}price\PYZus{}path}\PY{o}{.}\PY{n}{transpose}\PY{p}{(}\PY{p}{)}

    \PY{k}{return} \PY{p}{[}\PY{n}{share\PYZus{}price\PYZus{}path}\PY{p}{,}\PY{n}{firm\PYZus{}price\PYZus{}path}\PY{p}{]}  
\end{Verbatim}
\end{tcolorbox}

A Pearson test can be used to verify that the correlation is very close to the 0.2 target value. In Python this can be achieved as follows.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{5}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{}Testing share and firm price correlation}
\PY{n}{sample\PYZus{}size} \PY{o}{=} \PY{l+m+mi}{20}
\PY{n}{test} \PY{o}{=} \PY{n}{generate\PYZus{}share\PYZus{}and\PYZus{}firm\PYZus{}price}\PY{p}{(}\PY{n}{S0}\PY{p}{,} \PY{n}{v\PYZus{}0}\PY{p}{,} \PY{n}{risk\PYZus{}free}\PY{p}{,} \PY{n}{sigma}\PY{p}{,} \PY{n}{sigma\PYZus{}firm}\PY{p}{,} \PY{n}{corr}\PY{p}{,} \PY{n}{T}\PY{p}{,} \PY{n}{sample\PYZus{}size}\PY{p}{,} \PY{n}{timesteps} \PY{o}{=} \PY{l+m+mi}{20000}\PY{p}{)}

\PY{n}{share\PYZus{}ret} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{log}\PY{p}{(}\PY{n}{test}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{)}

\PY{k}{for} \PY{n}{i} \PY{o+ow}{in} \PY{n+nb}{range}\PY{p}{(}\PY{n}{sample\PYZus{}size}\PY{p}{)}\PY{p}{:}
    \PY{n}{test}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{sharelog}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{log}\PY{p}{(}\PY{n}{test}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{)}
    \PY{n}{test}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{firmlog}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{log}\PY{p}{(}\PY{n}{test}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{[}\PY{n}{i}\PY{p}{]}\PY{p}{)}
    \PY{n+nb}{print}\PY{p}{(}\PY{n}{stats}\PY{o}{.}\PY{n}{pearsonr}\PY{p}{(}\PY{n}{test}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{sharelog}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{diff}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{dropna}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{test}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{[}\PY{l+s+s1}{\PYZsq{}}\PY{l+s+s1}{firmlog}\PY{l+s+s1}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{diff}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{dropna}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

    \begin{Verbatim}[commandchars=\\\{\}]
(0.2035028479587627, 6.07336614624919e-186)
(0.1986599648970965, 4.04999708397334e-177)
(0.2016759847597355, 1.3775160789695309e-182)
(0.20364055709605317, 3.381815103189066e-186)
(0.19385401550478024, 1.3726860317349184e-168)
(0.19892212605701942, 1.3666646645127986e-177)
(0.18344242259502544, 7.064732561303709e-151)
(0.20519864966486748, 4.356484560250648e-189)
(0.193927020114363, 1.022531700201078e-168)
(0.19751037260327703, 4.659785935918803e-175)
(0.20393895669991588, 9.495454795149592e-187)
(0.20458792743312265, 5.953557128978809e-188)
(0.2042028358117276, 3.0829601588859696e-187)
(0.19512700379639908, 7.943982338391081e-171)
(0.19603082201925184, 2.003660281095667e-172)
(0.19820752618175266, 2.630853001655849e-176)
(0.20299979761482267, 5.137198920374959e-185)
(0.1966663418793928, 1.4903037901828931e-173)
(0.20319824398313788, 2.214204369119104e-185)
(0.20765530182748704, 1.0810080941108714e-193)
    \end{Verbatim}

The payoff calculation for the up-and-out barrier option is defined as follows.

    \begin{tcolorbox}[breakable, size=fbox, boxrule=1pt, pad at break*=1mm,colback=cellbackground, colframe=cellborder]
\prompt{In}{incolor}{11}{\boxspacing}
\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{\PYZsh{} define payoff for up\PYZhy{}and\PYZhy{}out call option}
\PY{k}{def} \PY{n+nf}{payoff}\PY{p}{(}\PY{n}{S\PYZus{}t}\PY{p}{,} \PY{n}{K}\PY{p}{,} \PY{n}{L}\PY{p}{)}\PY{p}{:}
    \PY{n}{stopped\PYZus{}S} \PY{o}{=} \PY{n}{S\PYZus{}t}\PY{o}{.}\PY{n}{iloc}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{o}{.}\PY{n}{where}\PY{p}{(}\PY{p}{(}\PY{n}{S\PYZus{}t} \PY{o}{\PYZlt{}} \PY{n}{L}\PY{p}{)}\PY{o}{.}\PY{n}{all}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{)}
    \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{maximum}\PY{p}{(}\PY{n}{stopped\PYZus{}S} \PY{o}{\PYZhy{}} \PY{n}{K}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{)}\PY{o}{.}\PY{n}{to\PYZus{}numpy}\PY{p}{(}\PY{p}{)}

\PY{n}{payoff}\PY{p}{(}\PY{n}{share\PYZus{}price\PYZus{}12\PYZus{}months}\PY{p}{,} \PY{n}{K}\PY{p}{,} \PY{n}{L}\PY{p}{)}
\end{Verbatim}
\end{tcolorbox}

The simulated payoff paths over twelve months are plotted below, with an arbitrary starting value of 200 for the firm value.

\begin{figure}
    \centering
    \caption{Simulated Payoff Paths over 12 Months}
    \includegraphics{images/all-paths.png}
\end{figure}

\section{Monte Carlo Estimations
}

\subsection{Default-Free Value Estimation}
\begin{figure}
    \centering
    \caption{Simulated Payoff Paths over 12 Months}
    \includegraphics{images/default-free.png}
\end{figure}

\subsection{Credit Valuation Adjustment Estimation}

@@

% the second parameter MMMMM should be as long as the longest label you use, in my case Smoller -- if you use % numbers only, use 99
% use \cite{refname} to refer to bibliography item \bibitem{refname} 
% LaTeX assigns a number, unless you use \bibitem[Name]{refname} -- in this case
% LaTeX prints Name when you use \cite{refname}
\begin{thebibliography}{MMMMM} 
\bibitem{BA1} Blanchard, Arnaud. "The two-factor Hull-White model: pricing and calibration of interest rates derivatives." KTH Royal Insitute of Technology (2014).

>>>>

@@

\section{Introduction
}
A European up-and-out call option is type of a Barrier Option, these options are consider path-dependant because barrier option's payoff is based on the underlying stock's price path.  The payoff of an up-and-out option at maturity $T$ is given by,

\[
Payoff_T= (S_T - K)^+ given, max_{t \in [0,T]} S_t < L
\]
Where, where $K$ is the strike of the option, $L$ is the barrier level, and $S_t$ is the share price at time $t$.  This is a type of call option whose payoff is reduced to 0 if the share price crosses the barrier level. \\

The seminal work of Merton \cite{Merton1} pioneered the formula for pricing barrier options.  This led pricing formulas under the geometric Brownian motion (GBM) framework for one-asset barrier options by Rich \cite{RD1} and multi-asset barrier options by Wong and Kwok \cite{WK1}. Although the return dynamics of underlying shares are not sufficiently well described by the GBM process proposed by Black and Scholes \cite{BS1}, we are going to assume both the stock and counterparty firm values follow GBM  with constant drift and volatilities and default only occurs at maturity.

\section{Lifetime simulations of the option. 
}
\section{Monte Carlo estimates of both the default-free value of the option and the Credit Valuation Adjustment (CVA).
}

@@

% the second parameter MMMMM should be as long as the longest label you use, in my case Smoller -- if you use % numbers only, use 99
% use \cite{refname} to refer to bibliography item \bibitem{refname} 
% LaTeX assigns a number, unless you use \bibitem[Name]{refname} -- in this case
% LaTeX prints Name when you use \cite{refname}
\begin{thebibliography}{MMMMM} 
\bibitem{Merton1} Merton, Robert C. "Theory of rational option pricing." The Bell Journal of economics and management science (1973): 141-183.
\bibitem{RD1} Rich, Don R. "The mathematical foundations of barrier option-pricing theory." Advances in futures and options research 7 (1994).
\bibitem{WK1} Wong, Hoi Ying, and Yue‐Kuen Kwok. "Multi-asset barrier options and occupation time derivatives." Applied Mathematical Finance 10.3 (2003): 245-266.
\bibitem{BS1}Black, Fisher, and Myron Scholes. "The pricing and Corportate Liabilities." Journal of Political Economy 81 (1973).
\bibitem{You1} Yousuf, M. "A fourth-order smoothing scheme for pricing barrier options under stochastic volatility." International Journal of Computer Mathematics 86.6 (2009): 1054-1067.
